---
title: "Generate synthetic nucleosome maps"
author: Rawane Samb, Astrid Louise Deschenes, Pascal Belleau and Arnaud Droit
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  BiocStyle::html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Generate synthetic nucleosome maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignettePackage{consensusSeekeR}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

## Licensing

This package and the underlying `r Rpackage("nucleoSim")` code are 
distributed under the Artistic license 2.0. You are free to use and 
redistribute this software.


## Loading nucleoSim package

As with any R package, the `r Rpackage("nucleoSim")` package should 
first be loaded with the following command:

```{r loadingPackage, warning=FALSE, message=FALSE}
library(nucleoSim)
```


## Session info

Here is the output of `sessionInfo()` on the system on which this document was 
compiled:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
